---
title: "kmeans clustering"
author: "Group 3"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library( tm )
library( plyr )
library(textstem)
library( cld3 )
library( ggplot2 )
library( cluster )

set.seed(1234)
```


## Load the data

Load the tweet data into a data frame for processing and show a summary of the data.

```{r tweets}
setwd("../data/")
tweets <- read.csv( "tweets_prepped.csv", header=TRUE, sep=",", as.is=FALSE )
str(tweets)

```

## Create a Term Document Matrix for processing

Create a term document matrix.

```{r tdm}
corpus <- Corpus(VectorSource(tweets$tweet))
corpusCopy <- corpus
tdm <- TermDocumentMatrix( corpus )
inspect(tdm)
```

Create a document term matrix.

```{r tmd}
tmd <- DocumentTermMatrix( corpus )
inspect(tmd)
findFreqTerms(tmd, lowfreq=400)
```

Graph term frequency.

```{r term frquency}
termFrequency <- colSums(as.matrix(tmd))
termFrequency <- subset(termFrequency, termFrequency>=400)
df <- data.frame(term=names(termFrequency), freq=termFrequency)
ggplot(df, aes(x=term, y=freq)) + geom_bar(stat="identity") + xlab("Terms") + ylab("Count") + coord_flip()

```

Generate k-means cluster.

```{r k-means}
tmd2 <- removeSparseTerms(tmd, sparse=0.95)
m2 <- as.matrix(tmd2)
#k <- 4
#k <- 8
k <- 16
kmeansResult <- kmeans(m2, k)
round(kmeansResult$centers, digits=3)
```

Look at top words in each cluster.

```{r top-cluster-words}
for (i in 1:k) {
   cat(paste("cluster ", i, ":  ", sep=""))
   s <- sort(kmeansResult$centers[i,], decreasing=T)
   cat(names(s)[1:10], "\n")
}
```

Look at top words in each cluster.

```{r plot-cluster}
clusplot(tweets, kmeansResult$cluster, color=TRUE, shade=TRUE, labels=1, lines=1)
```

Compare cluster predictions with actual comparisons.

```{r table-cluster}
table(tweets$label, kmeansResult$cluster)
```
