---
title: "Naive Bayes Model of Tweets_Project"
author: "Chukuma I."
date: "2022-08-06"
output: html_document
---
# Naive Bayes Model #

Install and load necessary packages and libraries for project

```{r}
library(e1071)
library(tidyverse)
library(ROCR)
library(dplyr)
library(stringr)
library(tidytext)
library(forcats)
library(textdata)
library("tm")
library("arules")
```


Set working directory and read csv file
Convert csv file into dataframe
Review stucture and view some rows
```{r}
setwd("F:/Project")
tweets_naive <- read.csv("enriched_tweet_data.csv", sep =",", header = TRUE)
str(tweets_naive)
glimpse(tweets_naive)
```

Create a corpus from the data frame of tweet
Convert corpus into a document term matrix
```{r}
# Create the corpus
tweets_naive_corpus <- Corpus(VectorSource(tweets_naive$tweet))

# create the DTM
tweets_naive_dtm <- DocumentTermMatrix(tweets_naive_corpus)

# cast the DTM to a useful format
tweets_naive_dtm <- as.matrix(tweets_naive_dtm)
```



Determine the median length of the corpus words and denote those which are below the median or below
```{r}
# Cast the link fields to an integer
tweets_naive_1 <- tweets_naive %>% mutate(link = as.integer(link)) %>% mutate(ssl_link = as.integer(ssl_link))


# Lets add a check to see if the tweet is larger or smaller than the median.
median_length <- median((tweets_naive_1 %>% mutate(nchar = nchar(tweet)))$nchar)
tweets_naive_1 <- tweets_naive %>% mutate(larger_than_median = nchar(tweet) > median_length) %>% mutate(larger_than_median = as.integer(larger_than_median))
```

Performing further preprocessing (transformation) on the data frame
```{r}
tweets_naive_1$anger<-ifelse(tweets_naive_1$anger>0, 1, tweets_naive_1$anger)
tweets_naive_1$anticipation<-ifelse(tweets_naive_1$anticipation>0, 1, tweets_naive_1$anticipation)
tweets_naive_1$disgust<-ifelse(tweets_naive_1$disgust>0, 1, tweets_naive_1$disgust)
tweets_naive_1$fear<-ifelse(tweets_naive_1$fear>0, 1, tweets_naive_1$fear)
tweets_naive_1$joy<-ifelse(tweets_naive_1$joy>0, 1, tweets_naive_1$joy)
tweets_naive_1$negative<-ifelse(tweets_naive_1$negative>0, 1, tweets_naive_1$negative)
tweets_naive_1$positive<-ifelse(tweets_naive_1$positive>0, 1, tweets_naive_1$positive)
tweets_naive_1$sadness<-ifelse(tweets_naive_1$sadness>0, 1, tweets_naive_1$sadness)
tweets_naive_1$surprise<-ifelse(tweets_naive_1$surprise>0, 1, tweets_naive_1$surprise)
tweets_naive_1$trust<-ifelse(tweets_naive_1$trust>0, 1, tweets_naive_1$trust)
tweets_naive_1$sentiment_afinn<-discretize(tweets_naive_1$sentiment_afinn, method="frequency", breaks = 4)

```



Further preprocessing requires removing some columns
```{r}
tweets_naive_1$ssl_link <- NULL
tweets_naive_1$tweet_orig<-NULL
tweets_naive_1$length<-NULL
tweets_naive_1$tweet<-NULL
tweets_naive_1$hashtag<-NULL
tweets_naive_1$sentiment_afinn<-NULL
tweets_naive_1$http<-NULL #Contains too many missing values

```

A view of the data frame
```{r}
glimpse (tweets_naive_1)
```


## Training ##

Set the seed value to ensure that result is reproducible
Then divide dataframe in ratio of 70:30 for train and test
```{r}

set.seed(1234)
# divide the data into training and test
ind <- sample(2, nrow(tweets_naive_1), replace = TRUE, prob = c(0.7, 0.3))
train.data <- tweets_naive_1 [ind == 1, ]
test.data <- tweets_naive_1 [ind == 2, ]
```

Using naive bayes model
label is the dependent variable
```{r}

model<-naiveBayes(label~., train.data)
model
prop.table(table(train.data$label)) #Returns the same output as A-priori probabilities in the model  
```

### Model evaluation on Training data ###

 Evaluate the model on a training set
```{r}
# This command returns the predicted probabilities for each class
predict(model, train.data, type="raw")
```

This command returns the predicted class
```{r}
predict(model, train.data, type="class") 

```

```{r}
table(predict(model, train.data, type="class"), train.data$label, dnn=c("predicted", "actual"))
```


classification accuracy for the training data
```{r}
mean(predict (model, train.data, type="class")== train.data$label)
```

 Classification error for the training data
```{r}
mean(predict (model, train.data, type="class")!= train.data$label)
```

### Model evaluation on Test data ###

Evaluate model on test data
```{r}
predict(model, test.data, type="raw")
```

This command returns the predicted probabilities for each class
```{r}
predict(model, test.data, type="class")
```

This command returns the predicted class
```{r}
table(predict(model, test.data, type="class"), test.data$label, dnn=c("predicted", "actual"))
```

classification accuracy for the training data
```{r}
mean(predict (model, test.data, type="class")== test.data$label)
```

Classification error for the training data
```{r}
mean(predict (model, test.data, type="class")!= test.data$label)
```


```{r}
mypredictions<-predict(model, test.data, type="class")
ROCRpred <- prediction(as.numeric(mypredictions)-1, test.data$label)
ROCRperf <- performance(ROCRpred, 'tpr','fpr')
plot(ROCRperf, colorize = TRUE, text.adj = c(-0.2,1.2), lwd=2)
abline(a=0,b=1,lwd=2,lty=2,col="gray")
```

## Mosaic Plot ##
mosaic plot is a special type of stacked bar chart that shows percentages of data in groups

```{r}
mosaicplot(table(predict(model, test.data), test.data$label), shade = TRUE, main = "Predicted vs. Actual LABEL")
```




sessionInfo()