---
title: "Neural-network (nnet) Tweet Analysis"
author: "Group 3"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library( nnet )
library( ggplot2 )
library( dplyr )
library ("NeuralNetTools")
require(caret)
require(e1071)
library("devtools")
library("ROCR")

set.seed(1234)
```


## Load the data

Load the tweet data into a data frame for processing and show a summary of the data.

```{r load}
setwd("../data/")
tweets <- read.csv( "enriched_tweet_data.csv", header=TRUE, sep=",", as.is=FALSE )
#tweets <- read.csv( "enriched_tweet_data_1.csv", header=TRUE, sep=",", as.is=FALSE )
#tweets <- read.csv( "enriched_tweet_data_2.csv", header=TRUE, sep=",", as.is=FALSE )
#tweets <- read.csv( "enriched_tweet_data_3.csv", header=TRUE, sep=",", as.is=FALSE )
#tweets <- read.csv( "enriched_tweet_data_4.csv", header=TRUE, sep=",", as.is=FALSE )
#tweets <- read.csv( "enriched_tweet_data_5.csv", header=TRUE, sep=",", as.is=FALSE )
#tweets <- read.csv( "enriched_tweet_data_6.csv", header=TRUE, sep=",", as.is=FALSE )
#tweets <- read.csv( "enriched_tweet_data_7.csv", header=TRUE, sep=",", as.is=FALSE )
#tweets <- read.csv( "enriched_tweet_data_8.csv", header=TRUE, sep=",", as.is=FALSE )
#tweets <- read.csv( "enriched_tweet_data_9.csv", header=TRUE, sep=",", as.is=FALSE )
str(tweets)

```

## Prep the data

Make any formating decisions necessary to the data before generation of the model.

```{r prep}
#tweets$id <- NULL
tweets$http <- as.factor(tweets$http)
tweets$label_factor <- ifelse(tweets$label == 'real', 0, 1) 
tweets$label_factor <- as.factor( tweets$label_factor )
tweets$length <- scale(tweets$length)
tweets$sentiment_afinn <- scale(tweets$sentiment_afinn)
tweets$anger <- scale(tweets$anger)
tweets$anticipation <- scale(tweets$anticipation)
tweets$disgust <- scale(tweets$disgust)
tweets$fear <- scale(tweets$fear)
tweets$joy <- scale(tweets$joy)
tweets$negative <- scale(tweets$negative)
tweets$positive <- scale(tweets$positive)
tweets$sadness <- scale(tweets$sadness)
tweets$surprise <- scale(tweets$surprise)
tweets$trust <- scale(tweets$trust)
tweets$tweet <- NULL
tweets$tweet_orig <- NULL
tweets$hashtag <- NULL
str(tweets)
summary(tweets)

```

## Create the model

Create the test and training datasets.

```{r datasets}
ind <- sample(2, nrow(tweets), replace = TRUE, prob = c(0.7, 0.3))
train.data <- tweets[ind == 1, ]
test.data <- tweets[ind == 2, ]
```

Create the neural network model.

```{r model}
nn <- nnet( label_factor~link+length+sentiment_afinn+anger+anticipation+disgust+fear+joy+negative+positive+sadness+surprise+trust, data = train.data, size=10, maxit=2000 )

# explore the model
names(nn)
summary(nn)  
```

## Evaluate the model

Visualize the neural net model

```{r plots}
plotnet(nn)
garson(nn)  
olden(nn)
```

Confusion matrices.

```{r tables}
mypredict<-predict(nn) # training
mypredict<-apply(mypredict, c(1), round)            
confusionMatrix(table(mypredict, train.data$label_factor, dnn=c("Predicted", "Actual")))

testPred <- predict(nn, test.data) #test
testPred<-apply(testPred, c(1), round)
confusionMatrix(table(testPred, test.data$label_factor, dnn=c("Predicted", "Actual")))

```

ROC Curve to evaluate the model.

```{r eval-curves}
pred <- prediction( testPred, test.data$label_factor )
perf_sn_sp = performance(pred, "sens", "spec")
plot(perf_sn_sp)

roc = performance(pred,"tpr","fpr")
plot(roc, colorize = T, lwd = 2)
abline(a = 0, b = 1) 

auc = performance(pred, measure = "auc")
print( auc@y.values )

```


```{r total_file}
# only run when evaluating the entirety of the dataset against a model trained with a cluster
runExtended <- FALSE
if ( runExtended ) {
  # run the new model on the full data set
  setwd("../data/")
  te <- read.csv( "enriched_tweet_data.csv", header=TRUE, sep=",", as.is=FALSE )
  te$id <- NULL
  te$http <- as.factor(te$http)
  te$label_factor <- ifelse(te$label == 'real', 0, 1)
  te$label_factor <- as.factor( te$label_factor )
  te$length <- scale(te$length)
  te$sentiment_afinn <- scale(te$sentiment_afinn)
  te$anger <- scale(te$anger)
  te$anticipation <- scale(te$anticipation)
  te$disgust <- scale(te$disgust)
  te$fear <- scale(te$fear)
  te$joy <- scale(te$joy)
  te$negative <- scale(te$negative)
  te$positive <- scale(te$positive)
  te$sadness <- scale(te$sadness)
  te$surprise <- scale(te$surprise)
  te$trust <- scale(te$trust)
  tweets$tweet <- NULL
  tweets$tweet_orig <- NULL
  tweets$hashtag <- NULL
  
  testPred <- predict(nn, te) #test
  testPred<-apply(testPred, c(1), round)
  confusionMatrix(table(testPred, te$label_factor), dnn=c("Predicted", "Actual"))
}

```