---
title: "Tweet Feature Extraction"
author: "Ada Lazuli"
date: '2022-07-17'
output: 
  html_document:
    toc: true
    toc_depth: 2
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, out.width = '100%', warning=FALSE, collapse=TRUE, class.source = 'fold-hide'}
library(tidyr)
library(dplyr)
library(ggplot2)
library(DT)
library(stringr)
library(tidytext)
library(forcats)
library(textdata)
```

# Data Loading

Load the prepped data compiled by Winfrey Johnson.

```{r,  out.width = '100%', warning=FALSE, message=FALSE}
df <- read.csv("../data/tweets_prepped.csv")
datatable(df)
```

# Data Enrichment 

To give the decision tree more features to work with, the data will be enriched to include:
1. A flag on whether HTTP is present
2. A numeric designation on overall sentiment using `afinn`
3. A other sentiment labels from the `nrc` corpus.

## Add Length

Use the `mutate` verb to add the length of the tweet to the dataframe

```{r,  out.width = '100%', warning=FALSE, message=FALSE}
df <- df %>% mutate(length = nchar(tweet))
```

## Flag HTTP

Use `mutate` to find the http keyword in the tweet text then convert that logical to an integer.

```{r,  out.width = '100%', warning=FALSE, message=FALSE}
df <- df %>% mutate(http = grepl("http", tweet)) %>% mutate(http = as.integer(http))
```

## Add Integer Sentiment

First, create a function that calculate the sentiment positivity value for one tweet using `afinn`. Then apply that function over all of the tweets in the dataframe.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
calc_afinn_sentiment <- function(df){
  df$sentiment_afinn <- 0
  for (i in 1 : nrow(df)){
    # extract the row
    temp <- df[i,]
    # un nest the row to join with afinn then calculate the average sentiment
    temp <- temp %>% unnest_tokens(word, tweet) %>% inner_join(get_sentiments("afinn"), by = "word") %>% group_by(label) %>% summarize(sentiment = mean(value))
    # sometimes the sentiment is not returned if the intersection of the sets is empty, so check for this below
    if (length(temp$sentiment) > 0){
      df[i,]$sentiment_afinn <- temp$sentiment[1] 
    }
  }
  return (df)
}
df <- calc_afinn_sentiment(df)
```


## Add Sentiement Labels

Similar to the previous section, create a function to add sentiment labels to each row of the dataframe using the `nrc` dataset.

```{r, out.width = '100%', warning=FALSE, message=FALSE}

find_nrc_sentiment <- function(df){
  # Add empty columns to the dataset for each label prior to calculation
  nrc_sentiments <- c("anger", "anticipation", "disgust",  "fear", "joy", "negative",  "positive", "sadness", "surprise",  "trust")
  for (x in 1:length(nrc_sentiments)){
    df[nrc_sentiments[x]] <- 0
  }
  
    for (i in 1 : nrow(df)){
    # extract the row
    temp <- df[i,]
    # un nest the row to join with nrc then count the occurrences of each sentiment
    temp <- temp %>% unnest_tokens(word, tweet) %>% inner_join(get_sentiments("nrc"), by = "word") %>% count(sentiment)
    # sometimes the sentiment is not returned if the intersection of the sets is empty, so check for this below
    if (nrow(temp) > 0){
      for (x in 1:nrow(temp)){
        df[i,][temp[x,]$sentiment] <- temp[x,]$n
      }
    }
  }
  return(df)
}
df <- find_nrc_sentiment(df)
```

## Save Enriched Dataset

Finally, save the enriched dataset as a CSV.

```{r}
write.csv(df, "../data/enriched_tweet_data.csv")
datatable(df)
```



























